Choosing TTC x Provides the Most Time Stability during Rush Hours

### 2024-09-21 outputs/paper/paper.qmd
created new tbls and other visualizations for data & result sections

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(lubridate)
library(zoo)
library(knitr)
library(ggplot2)
```

@@

bus_delay <- read.csv(here::here("inputs/data/cleaned_data/bus_delay.csv"))

@@

streetcar_delay <- read.csv(here::here("inputs/data/cleaned_data/streetcar_delay.csv"))

@@

subway_delay_code <- read.csv(here::here("inputs/data/cleaned_data/subway_delay_code.csv"))

@@

subway_delay <- read.csv(here::here("inputs/data/cleaned_data/subway_delay.csv"))

@@

```{r}
#| label: fig-coded
#| fig-cap: Visualization to show correlation between number of Calls Coded (represented by the red line) and Homeless Death Counts (represented by the blue line)
#| echo: false
#| message: false

combined_data <- read_csv(here::here("outputs/data/combined_data.csv"),
                          show_col_types = FALSE)

combined_data|> ggplot(mapping = aes(x = Month, y = Death_Count)) + 
  geom_point() + geom_smooth() + 
  geom_line(aes(x = Month, y = Calls_Coded_month/1000), color = "red") +
  labs(
    title = "Calls Coded & Homeless Death Count",
    x = "Time (in Month)",
    y = "Count"
  )
#still need to add a legend or label, but I haven't yep figured out how
```

@@

Present Results During Rush Hours & at Off-Peak Hours
```{r}
#| label: fig-referred
#| fig-cap: Visualization to show correlation between number of callers referred to shelter (represented by the purple line) and Homeless Death Counts (represented by the blue line)
#| echo: false
#| message: false

combined_data <- read_csv(here::here("outputs/data/combined_data.csv"),
                          show_col_types = FALSE)

combined_data|> ggplot(mapping = aes(x = Month, y = Death_Count)) + 
  geom_point() + geom_smooth() + 
  geom_line(aes(x = Month, y = Referral_to_Shelter_month/100), color = "purple") +
  labs(
    title = "Callers Referred to Shelter & Homeless Death",
    x = "Time (in Month)",
    y = "Count"
  )
#still need to add a legend or label, but I haven't yep figured out how
```

@@

combined_data <- read_csv(here::here("outputs/data/combined_data.csv"),
                          show_col_types = FALSE)

combined_data|> ggplot(mapping = aes(x = Month, y = Death_Count)) + 
  geom_point() + geom_smooth() + 
  geom_line(aes(x = Month, y = Information_Homelessness_and_Prevention_month/100), color = "darkgreen") +
  labs(
    title = "Callers given Info on Homeless Prevention vs Homeless Death",
    x = "Time (in Month)",
    y = "Count"
  )
#still need to add a legend or label, but I haven't yep figured out how
```

## Most Popular Delay Reasons
- 1 table for all three services

## Most Popular Delayed Vehicles
- 1 table for all three services

>>>>

```{r}
#| include: false
#| warning: false
#| message: false
# load librarys and datasets
library(tidyverse)
library(dplyr)
library(lubridate)
library(zoo)
library(knitr)
library(ggplot2)

bus_delay <- read.csv(here::here("inputs/data/cleaned_data/bus_delay.csv"))
streetcar_delay <- read.csv(here::here("inputs/data/cleaned_data/streetcar_delay.csv"))
subway_delay <- read.csv(here::here("inputs/data/cleaned_data/subway_delay.csv"))
subway_delay_code <- read.csv(here::here("inputs/data/cleaned_data/subway_delay_code.csv"))

```

@@

@@

@@

@@

@@

```{r}
#| label: fig-total-delay
#| fig-cap: Visualization to show the TTC service that has the highest number of delays in year 2023
#| echo: false
#| message: false
#construct appropriate table for visualization
service_type <- c("bus", "streetcar", "subway")
num_delay <- c(nrow(bus_delay), nrow(streetcar_delay), nrow(subway_delay))

delay_count <- data.frame(service_type, num_delay)

# Generate the bar plot
ggplot(delay_count, aes(x = service_type, y = num_delay, fill = service_type)) +
  geom_bar(stat = "identity") + 
  labs(title = "Number of Delays by Service Type", x = "Service Type", y = "Number of Delays", fill = "Service Type") +  # Label the axes
  theme_minimal()  # Apply a minimal theme
```

@@

Present Results During Rush Hours & at Off-Peak Hours

```{r}
#| echo: false
#| message: false

#filter out delays that occurred on weekdays
weekday_bus_delay <- bus_delay |>
  filter(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
weekday_streetcar_delay <- streetcar_delay |>
  filter(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
weekday_subway_delay <- subway_delay |>
  filter(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))

#filter out delays that occurred on weekends
weekend_bus_delay <- bus_delay |>
  filter(day %in% c("Saturday", "Sunday"))
weekend_streetcar_delay <- streetcar_delay |>
  filter(day %in% c("Saturday", "Sunday"))
weekend_subway_delay <- subway_delay |>
  filter(day %in% c("Saturday", "Sunday"))
```

```{r}
#| label: fig-weekday
#| fig-cap: Visualization to show the most delayed services on weekdays rush hours and off-peak hours
#| echo: false
#| message: false

```

```{r}
#| label: fig-weekend
#| fig-cap: Visualization to show the most delayed services on weekdays rush hours
#| echo: false
#| message: false

```

@@

```

## Most Popular Delay Reasons
- 1 table for all three services
- separate into rush & off-peak hours

## Most Popular Delayed Vehicles
- 1 table for all three services
- separate into rush & off-peak hours
